---
title: "p8105_hw2_fx2184"
author: "Fei"
date: "2022-09-29"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(readr)
library(readxl)
```

```{r}
subway = read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv")%>% janitor:: clean_names() %>% 
    mutate(
      entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE" ),
      entry = as.logical(entry)
    )
skimr::skim(subway)
```


* The dataset NYC Transit Subway Entrance and Exit Data contains `r nrow(subway)` x `r ncol(subway)` variables with name: `r names(subway)`. 

* The data describes the 11 different routes of NYC. For now, I only imported this data and change the variable type of entry to logical variable. 
* This data is not a tidy data due to redundancy.  

# How many distinct stations are there? 

```{r}
# How many distinct stations are there? 
distinct_stations = distinct(subway, line, station_name, .keep_all = TRUE)
# How many stations are ADA compliant?
ada = distinct(subway, line, station_name, .keep_all = TRUE) %>%  count(ada)
# What proportion of station entrances / exits without vending allow entrance?
proportion = nrow(filter(subway, vending == "NO", entry == "TRUE"))/nrow(filter(subway, vending == "NO"))
show(proportion)
#How many distinct stations serve the A train? 
subway_wide = subway %>%
  mutate(route8 = as.character(route8),
         route9 = as.character(route9),
         route10 = as.character(route10),
         route11 = as.character(route11)) %>%
  pivot_longer(route1:route11, names_to = "route_number", values_to = "route_name")

distinct_A = distinct(subway_wide, line, station_name, .keep_all = TRUE) %>% count(route_name == "A")

# Of the stations that serve the A train, how many are ADA compliant?
distinct_A_ada = distinct(subway_wide, line, station_name, .keep_all = TRUE) %>% count(ada & route_name == "A")

```
* Answer: there are `r nrow(distinct_stations)` distinct stations in this dataset.
* Answer: There are 84 stations are ADA compliant, and 381 stations are not ADA compliant. 
* Answer: The proportion of station entrances / exits without vending allow entrance is `proportion`.
* Answer: there are 60 distinct station serve the A train. 
* Answer: there are 17 distinct station serve A train and ada compliant. 


# Problem 2 

## Mr. Trash Wheel dataset

```{r}
Mr_Trash_Wheel = read_excel("~/Desktop/p8105_hw2_fx2184/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                         sheet = "Mr. Trash Wheel") %>% 
    janitor:: clean_names() %>% 
    select(-x15,-x16,-x17) %>%
    filter(dumpster !='Grand Total') %>% 
    drop_na(dumpster) %>% 
    rename("weight(tons)" = "weight_tons", "volume(cubic yards)" = "volume_cubic_yards") %>% 
    mutate(
      sports_balls = as.integer(round(sports_balls,0)))
```

## Professor Trash Wheel

```{r}
Pro_Trash_Wheel = read_excel("~/Desktop/p8105_hw2_fx2184/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                         sheet = "Professor Trash Wheel") %>% 
    janitor:: clean_names() %>% 
    drop_na(dumpster) %>% 
    mutate(sports_balls = as.integer(round(sports_balls,0)),
           Trash_wheel = c('Pro') )%>% 
  #add a new column called Trash_wheel in order to merge two datasets in the next step
    rename("weight(tons)" = "weight_tons", "volume(cubic yards)" = "volume_cubic_yards") 

```

## Merge Mr and Professor Trash Wheel 

```{r}
# add a new column Trash_wheel in Mr Trash Wheel in order to merge two datasets.
Mr_Trash_Wheel = mutate(Mr_Trash_Wheel, Trash_wheel = c('Mr'))
Pro_Trash_Wheel = mutate(Pro_Trash_Wheel,dumpster = as.character(dumpster))
# 
Mr_and_Pro = Pro_Trash_Wheel %>% full_join(Mr_Trash_Wheel)

#sum values in column weight(tons) where Trash_wheel is equal to 'Pro'
total_weight = sum(Mr_and_Pro[which(Mr_and_Pro$Trash_wheel=='Pro'), "weight(tons)"])

#sum values in column sports_balls where Trash_wheel is equal to 'Mr'
total_balls = sum(Mr_and_Pro[which(Mr_and_Pro$Trash_wheel=='Mr'& Mr_and_Pro$year == '2020'), "sports_balls"])

```

* The merged dataset Mr_and_Pro has `r nrow(Mr_and_Pro)` observations and `r ncol(Mr_and_Pro)` variables, they are `r names(Mr_and_Pro)`

* The total weight of trash collected by Professor Trash Wheel is `r total_weight` tons.  

* The total number of sports balls collected by Mr. Trash Wheel in 2020 is  `r total_balls`. 

# Problem 3 （fivethirtyeight datasets) 

# Dataset: pols-month 

```{r}
pols_month = read.csv("~/Desktop/p8105_hw2_fx2184/fivethirtyeight_datasets/pols-month.csv") 
pols_month = pols_month %>% 
  janitor:: clean_names() %>%
  mutate(pols_month, mon = as.Date(mon)) %>% 
  mutate(year = as.integer(format(mon, format = "%Y")),
         month = as.integer(format(mon, format = "%m")),
         day = as.integer(format(mon, format = "%d"))) %>% 
  mutate(month = month.abb[month]) %>% 
  mutate(president = ifelse(prez_gop == "0", "dem","gop")) %>% 
  select(-prez_dem, -prez_gop,-mon)
```

# Dataset: snp dataset

```{r}
snp = read.csv("~/Desktop/p8105_hw2_fx2184/fivethirtyeight_datasets/snp.csv") 
snp = snp %>% 
  janitor:: clean_names() %>%
  mutate(snp, date = as.Date(date)) %>% 
  mutate(month = as.integer(format(date, format = "%Y")),
         day = as.integer(format(date, format = "%m")),
         year = as.integer(format(date, format = "%d"))) %>% 
  mutate(month = month.abb[month]) %>% 
  arrange(desc(year),desc(month)) %>% 
  select(year, month, day,close,-date)

```

#  Merge 

```{r}
unemployment = read.csv("~/Desktop/p8105_hw2_fx2184/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemployment_rate") %>% 
  janitor:: clean_names()
 # mutate(unemployment, month = as.integer(month))
pols_snp = pols_month %>% left_join(snp,Joining, by = c("year", "month", "day"))
merge = pols_snp %>% left_join(unemployment,by = c("year", "month"))
```

* The `pols_month` dataset has `r nrow(pols_month)`observations with `r ncol(pols_month)` variables: ` r names(pols_month)`. The range of year is [`r pols_month %>%  pull (year) %>% range()`].

* The `snp` dataset has has `r nrow(snp)`observations with `r ncol(snp)` variables: ` r names(snp)`. The range of year is [`r snp %>%  pull (year) %>% range()`].

* The `unemployment` dataset has has `r nrow(unemployment)`observations with `r ncol(unemployment)` variables: ` r names(unemployment)`. The range of year is [`r unemployment %>%  pull (year) %>% range()`].

* First, using the left join to merge the snp dataset into pols_month, the new dataset has `r nrow(pols_snp)`observations and `r ncol(pols_snp)`variables,they are ` r names(pols_snp)`. The range of year is [`r pols_snp %>% pull(year) %>% range()`]. Then, merging the `unemployment` dataset to `pols_snp`, It contains `r nrow(merge)`observations and `r ncol(merge)` variables:` r names(merge)`. The range of year is [`r merge %>% pull(year) %>% range()`].


